Evaluating the Performance and Energy Efficiency of COSMO-ART, a Coupled Numerical Weather Forecast and Chemical Transport Model

### 2014-04-10 paper_enahpc14/paper.tex
adds latest modifications to the document
Manuel's changes

\author{Joseph~Charles  \and William~Sawyer  \and  Manuel~F.~Dolz \and
  Cristiano~Malossi}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
	%\\ Tel: +41  (0) 91  610 8216
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg
	\\ D-22.527 Hamburg, Germany
        %\\ Tel: +49 (0) 40 460094-404
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and C.~Malossi  
        \at IBM Zurich Research Laboratory 
        \\  CH-8000 Zurich, Switzerland
        %\\  Tel: +41 (0) 44 724 8616
        \\ \email{ACM@zurich.ibm.com}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  weather  service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is analysed  by a profiling study to  determine hotspots and
  identify critical paths.   Moreover, we describe measurement devices
  and  energy-aware   techniques  employed  to   evaluate  the  energy
  footprint of the considered application and to get detailed insights
  about power bottlenecks.  Our motivation is to improve corresponding
  code   sections  to  sustain   high  performance   while  minimizing
  energy-to-solution.   This  preliminary  work  sets  the  basis  for
  subsequent studies to tackle  challenges related to energy efficient
  high performance computing in the framework of the Exa2Green project
  (\url{http://exa2green.eu/}).

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark analysis \and COSMO-ART coupled model}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
\input{Introduction.tex}

\section{The COSMO-ART model system}
\label{sec:1}
\input{Sec1:COSMO-ART.tex}

\section{Power-performance measurement framework}
\label{sec:2}
\input{Sec2:Power-measurement.tex}

\section{Evaluation}
\label{sec:3}
\input{Sec3:Evaluation.tex}

\section{Related work}
\label{sec:4}
\input{Sec4:Related.tex}

\section{Conclusion}
\label{concl}
\input{Conclusion.tex}

@@

%\begin{figure*}
%  \includegraphics[width=0.4\textwidth]{Figs/earth.eps}
%  \caption{Please write your figure caption here}
%  \label{fig:1}
%\end{figure*}

%\begin{table}
%  \caption{Please write your table caption here}
%  \label{tab:1}
%  \begin{tabular}{lll}
%    \hline\noalign{\smallskip}
%    first & second & third  \\
%    \noalign{\smallskip}\hline\noalign{\smallskip}
%    number & number & number \\
%    number & number & number \\
%    \noalign{\smallskip}\hline
%  \end{tabular}
%\end{table}

\begin{acknowledgements}
This  research  was  supported   by  the  Exa2Green  research  project
co-funded  under the  EU  7th Framework  Program  Future and  Emerging
Technologies (FET) Proactive Initiative: Minimising Energy Consumption
of Computing to  the Limit (MINECC).  The authors  would like to thank
the  High  Performance  and  High  Productivity  Computing  Initiative
(\url{www.hp2c.ch}) for  results that will be  leveraged in subsequent
code refactoring.
\end{acknowledgements}

>>>>

\author{Joseph~Charles  \and William~Sawyer  \and  Manuel~F.~Dolz \and
  Sandra~Catal\'an}

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg (UHAM)
	\\ DE-22527 Hamburg, Germany
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and S.~Catal\'an
        \at Jaume I University of Castell\'on (UJI)
	\\ ES-12071 Castell\'on, Spain
        \\ \email{catalans@icc.uji.es}}

@@

\begin{abstract}
  In this paper we present  COSMO-ART, an extension of the operational
  weather  forecast  model  of   the  German  weather  service  (DWD),
  developed for  the evaluation of the interactions  of reactive gases
  and aerosol particles  with the state of atmosphere  at the regional
  scale. It  includes secondary aerosols,  directly emitted components
  like  soot,  mineral  dust,  sea  salt and  biological  material  as
  pollen. Processes such  as emissions, coagulation, condensation, dry
  deposition,  wet removal,  and sedimentation  of aerosols  are taken
  into account.   The overall performance  of this application  on HPC
  systems is analysed  by a profiling study to  determine hotspots and
  identify  critical  CPU paths.   Moreover,  we describe  measurement
  devices and energy-aware techniques  employed to evaluate the energy
  footprint of the considered application and to get detailed insights
  about power bottlenecks.  Our motivation is to improve corresponding
  code   sections  to  sustain   high  performance   while  minimizing
  energy-to-solution.   This  preliminary  work  sets  the  basis  for
  subsequent studies to tackle  challenges related to energy efficient
  high performance computing in the framework of the Exa2Green project
  (\url{http://exa2green.eu/}).

\keywords{High performance computing  \and Energy-aware computing \and
  Green computing  \and Numerical weather  prediction \and Atmospheric
  chemistry  \and  Aerosols  modelling  \and  Profiling  methods  \and
  Benchmark    analysis   \and    COSMO-ART    coupled   model    \and
  Time-to-solution \and Energy-to-solution \and Manycore}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{sec:1}
\input{Sec1:Introduction.tex}

\section{The COSMO-ART model system}
\label{sec:2}
\input{Sec2:COSMO-ART.tex}

\section{Experimental setup}
\label{sec:3}
\input{Sec3:Power-measurement.tex}

\section{Evaluation}
\label{sec:4}
\input{Sec4:Evaluation.tex}

\section{Related work}
\label{sec:5}
\input{Sec5:Related.tex}

\section{Conclusion}
\label{sec:6}
\label{Sec6:Conclusions.tex}
\input{Conclusion.tex}

@@

\begin{acknowledgements}
This  research  was  supported   by  the  Exa2Green  research  project
co-funded  under the  EU  7th Framework  Program  Future and  Emerging
Technologies (FET) Proactive Initiative: Minimising Energy Consumption
of Computing  to the Limit  (MINECC).  We also  gratefully acknowledge
the  High  Performance  and  High  Productivity  Computing  Initiative
(\url{www.hp2c.ch}) for  results that will be  leveraged in subsequent
code refactoring.
\end{acknowledgements}
